---
title: "All the Right Questions August 2026"
execute:
  echo: false
---

## Welcome!

Welcome to the 2nd edition of *All the Right Questions:* The new monthly newsletter for Church Army staff in which we think together about how we can get, share and use the data that we need to do the work that God has called us to. The format is still evolving, but I'm hoping to use this space to signpost resources, celebrate successes and remind you all that I exist and that I care about your professional relationship with data.

Thanks so much for joining me here. If you've got any questions, or if there's anything you'd like to see covered in a future edition of *ARQ*, please let me know! Drop me an email (I'm david.lovell), or just come and say hi to me at my desk.

This newsletter isn't an official Church Army publication and as such it doesn't represent the views of Church Army. Just me, Data Dave.

## This week:

-   [Really Good Survey Training in October](#sec-promotion): I'm very excited about it
-   [The lesson](#sec-lesson): Alex Dobson has psychic powers
-   [A Small Idea](#sec-big-idea): The grammar of data collection
-   [An incredible achivement](#sec-achievement): Printing custom QR-codes onto sticky-labels
-   [The puzzle](#sec-puzzle): Not really a puzzle at all - test yourself for psychic powers and get featured in the next episode of *All the Right Questions*!

## Really Good Survey Training in October! {#sec-promotion}

## A small idea: The grammar of data collection {#sec-big-idea}

## Something I'm proud of: Customised QR-codes on address labels! {#sec-achievement}

## The lesson: {#sec-lesson}

Last month, we asked a bold and timely question. Are any of our colleagues psychic? This month, we're going to get some answers. 

For those of you who are just joining us, here is a brief recap:

- We got beloved colleagues (hereafter 'subjects') to guess 10 randomly drawn cards
- We assigned a 'success score' based on the improbability of correct guesses
- We postulated the existence of a 'psychic score' to describe a subject's relative psychic power

This month, we'll look at each subject's success score and ask the question 'what level of psychic power best explains this result?'

Let's start with our favourite and most potentially psychic subject: Alex Dobson. Alex got an astounding success score of 82.3 - more than double the expected average. The task before us is to establish the link between the measure of Alex's success and the measure of his psychic ability. The tool that we will use for that task is simulation. We'll ask the computer to pretend that it has a psychic score of 0, and then draw 10 cards and calculate its total success score. We will ask it to do this 4 million times and to give us 4 million scores. Then we will do the same thing again for a psychic score of 0.01, then 0.02, all the way up to a psychic score of 1. That's a total of 40 billion cards flipped, which takes the computer a few minutes to work out. 

Next, we take Alex's success score of 82.3, and compare it with the results of our simulation. For each interval of psychic score, what proportion of our simulated results are equal or higher to 82.3? Here is what that looks like:


```{r psychic_business}
#| cache: false
#| fig-cap: "Purple is the colour of psychic"
library(purrr)
library(here)
library(ggplot2)
library(dplyr)
library(scales)

set.seed(pi)

flip_cards <- readRDS(here("2026", "data", "flip-cards-function.RDS"))

psychic_scores <- seq(0, 1, 0.01)
success_scores <- map(psychic_scores, \(x) flip_cards(10, 4e6, x))


success_ratio <- function(x, threshold){
  sum(x >= threshold) / length(x)
}

scores <- list(
  anon_1 = 26/11 * 10,
  JenniK = 17.333,
  MagicIanthemagician = 190.666/52 * 10
)

score_prob <- map_dbl(success_scores, success_ratio, 82.3)
score_prob_anon_1 <- map_dbl(success_scores, success_ratio, scores$anon_1)
score_prob_JenniK <- map_dbl(success_scores, success_ratio, scores$JenniK)
score_prob_magician <- map_dbl(success_scores, success_ratio, scores$MagicIanthemagician)
score_prob_dan <- map_dbl(success_scores, success_ratio, 4.33)

len <- length(score_prob)

scores <- tibble(
  
  subject = 
    rep(
    c("Alex", "Anon", "JenniK", "MagicIanthemagician", "Dan Palmer"),
    each = len
  ),
  
  score_prob = c(score_prob, score_prob_anon_1, score_prob_JenniK, score_prob_magician, score_prob_dan)
)



results <-
  group_by(scores, subject) |> 
  mutate(psych_score = psychic_scores,
         norm_score_prob = (score_prob - min(score_prob)) / max(score_prob),
         norm_prob_diff = norm_score_prob - lag(norm_score_prob, default = 0)) 

results |> 
  filter(subject == "Alex") |> 
  
  ggplot(aes(x = psych_score, y = score_prob)) +
  
  geom_area(fill = "blue", colour = "black", alpha = 0.5) +
  
  labs(x = "Psychic score", y = "% beating 82.3", title = "Subject: Alex Dobson") +
  
  scale_y_continuous(labels = label_percent(1)) +
  
  theme_minimal()
```




```{r plotto}


ggplot(results) + 
  
  geom_area(aes(x = psych_score, y = norm_prob_diff, fill = subject),
            alpha = 0.5, position = "dodge", colour = "black") + 
  
  theme_minimal() +
  
  labs(
    x = "Psychic score",
    y = "Probability density"
  )

```

## The Puzzle {#sec-puzzle}

Not actually a puzzle at all. It's your turn to test yourself for psychic powers, using this [clever widget](https://church-army.shinyapps.io/psychic-tester/) I've built. Tune in next month to find out whether or not you are psychic!
