---
title: "All the Right Questions August 2026"
execute:
  echo: false
---

## Welcome!

Welcome to the 2nd edition of *All the Right Questions:* The new monthly Church Army newsletter about all things data. As always nothing here represents Church Army views - just mine.

Thanks so much for joining me here. If you've got any questions, or if there's anything you'd like to see covered in a future edition of *ARQ*, please let me know! Drop me an email (I'm david.lovell), or just come and say hi to me at my desk.

## This week:

-   [The Big Idea](#sec-big-idea): Is Church Army a Device?
-   [The lesson](#sec-lesson): Exactly how psychic are we?
-   [Really Good Survey Training in October](#sec-promotion): Spaces are filling
-   [An incredible achivement](#sec-achievement): Printing custom QR-codes onto sticky-labels
-   [The puzzle](#sec-puzzle): Not really a puzzle at all - test yourself for psychic powers and get featured in the next episode of *All the Right Questions*!

## The Big Idea: Is Church Army a Device? {#sec-big-idea}

Did you know that 'focus' is the Latin word for 'fireplace'?

For centuries, the hearth was the centre of the home. It was the warmth around which a family gathered, and the light by which they saw one another every evening. Pre-modern people did not take warmth and light for granted, but carefully maintained them through an unwavering pattern of familial responsibility consisting of things like the chopping of wood, the shaving of kindling, the daily construction, conflagration and cleansing of the fire itself. The hearth was not merely a device for producing heat, but an object through which a family became entangled in mutual responsibility, the passing of the day and the enjoyment of one another's company. 

I'm currently reading Albert Borgmann's _Technology and the Character of Contemporary Life_ (1984). It's a bit of a challenging book but very enjoyable and pertinent. The big idea of the book is that modern technological devices are starkly different from the tools that preceded them: The radio and the violin, for instance, are both designed to provide music, but a radio is not a violin. For Borgmann, the most important difference is that the things modern devices 'produce' have almost nothing to do with the way that they work - the radio 'produces' music by means of electrical transistors, but the relationship between the music and the transistors remains a mystery. Through radios and streaming and noise-cancelling headphones I am promised all the music I want, all to myself, whenever I want it, without any effort. But the hidden cost of the promise is, in short, the decline of the local folk music scene. Where once music was understood to be something local, skillful, shared and costly, it has now become something place-less, effortless, insular and cheap.

Why am I prattling on about this? I don't know if I've done it justice, but I've found Borgmann's 'device paradigm' to be an exciting and illuminating lens through which to view my relationships with the modern world, partciularly in regard to its promise of predictable results, repeatable experiences and satisfaction without toil. And it's got me to thinking - with my Data Insights hat on - what might be the benefits of reflecting on the modern charity as a kind of device? And I think that there are many.

For supporters, the 'charity device' delivers a commodified experience of participation without effort. You open an email and watch a video. You read a well-crafted paragraph in which you are thanked for your involvement in the work. You are persuaded that you are connected to these places and people, and that you are making impact happen. This participation is earnest and in some measure costly, but it is almost exclusively mediated through the manufacture of digital goods via digital devices. One device creates a digital artifact, another device 'delivers' it to a person hundreds of miles away, and a further device (we hope) 'sends' an amount of money to fund the work. These are the constituent parts of a supporter engagement device, which enables a scalable approach to fundraising that is unconstrained by geography. The cost is the inherent reduction of partenership to an abstract exchange of commodities.

This is true not only of our supporters, but also of some us who work within the Mission Support Team. I have personally spent almost a decade doing working adjacent to Centres of Mission, but in all that time I have never actually set foot in one. I have gathered data and summarised findings; I have never stacked chairs or refilled the pantry. I've compiled sprawling matrices describing our work, but I have never opened up on a frosty morning in February, or had my day interrupted by the needs of someone coming through the door in need of help. Our evangelists are deeply entangled in these places, they know them as the ancient family new its hearth. I know these places only through my devices - through the commodities that my machines produce. I turn these places off at 5 o'clock.

I know that this is a very selective way of thinking about our work. I know that lots of us *have* visited centres of mission, or attended (and planned!) residential Gatherings, or fostered warm and personal relationships with supporters and churches over the phone. But I still believe that the device paradigm can help us, particularly because of the ways it accords with perceived differences in the outlook of front-line workers and the mission support team, and especially in relation to how we measure 'impact'. I've taken enough of your time, so I'll summarise these briefly:

- For the distant and disembodied observer, 'impact' is treated an abstract commodity produced by the centre-of-mission-device. Totaling up measures of impact across diverse contexts belies an understanding that these devices can be thought of as equivalent because they each produce the same commodity. For the observer, the 'means' are secondary to the 'ends'. For the evangelist, the two can not be meaninfully separated.
- In my time as a researcher, it was not uncommon to hear evangelists 'push back' against a perceived expectation of consistent and predictable results. Such results are normally expected from devices, and are expected in general because of the prevalence of the device paradigm. Evangelists were essentially concerned that we understood their ministry as a device, and that we did not care to understand how it worked.
- More generally, there was an espoused sense of 'distance' on all sides - a displacement from one another's contexts leading to a lack of mutual understanding. This distance is a necessary consequence of devices we use, which enable us to speak to a person without really being with them.

I hope that those observations are persuasive and give you something to chew on. But I will leave you with one last idea:

Borgmann concludes his book by advocating wise living in the modern world through what he calls 'folk practices'. These are things like the hearth, which integrate the mind and the body with the physical, natural, social and temporal layers of reality. Things like linocutting, or playing the violin, or playing Magic the Gathering with the boys on a Tuesday night. These things bring us into contact with people, with place and with time. They difficult and unpredictable, and invite repeated failure. 

Our evangelists are folk practitioners. This is - if you will excuse the crudity of the term - their 'unique selling point'. In a world that is increasingly dispersed and predictable, they embody the gospel and enmesh themselves in the mess and magnificence of a particular place. They give themselves to people, to place and to time. Their work is difficult and unpredictable, and it invites repeated failure. But it is real, it is consistent and it is patiently embodied. Perhaps the difficulty we have had in summarising the impact of such work is a vital feature of its resistance to the imposition of the paradigm of the modern device.


## Really Good Survey Training in October! {#sec-promotion}

## Something I'm proud of: Customised QR-codes on address labels! {#sec-achievement}

## The lesson: {#sec-lesson}

Last month, we asked a bold and timely question. Are any of our colleagues psychic? This month, we're going to get some answers. 

For those of you who are just joining us, here is a brief recap:

- We got beloved colleagues (hereafter 'subjects') to guess 10 randomly drawn cards
- We assigned a 'success score' based on the improbability of correct guesses
- We postulated the existence of a 'psychic score' to describe a subject's relative psychic power

This month, we'll look at each subject's success score and ask the question 'what level of psychic power best explains this result?'

Let's start with our favourite and most potentially psychic subject: Alex Dobson. Alex got an astounding success score of 82.3 - more than double the expected average. The task before us is to establish the link between the measure of Alex's success and the measure of his psychic ability. The tool that we will use for that task is simulation. We'll ask the computer to pretend that it has a psychic score of 0, and then draw 10 cards and calculate its total success score. We will ask it to do this 4 million times and to give us 4 million scores. Then we will do the same thing again for a psychic score of 0.01, then 0.02, all the way up to a psychic score of 1. That's a total of 40 billion cards flipped, which takes the computer a few minutes to work out. 

Next, we take Alex's success score of 82.3, and compare it with the results of our simulation. For each interval of psychic score, what proportion of our simulated results are equal or higher to 82.3? Here is what that looks like:


```{r psychic_business}
#| cache: false
#| fig-cap: "Purple is the colour of psychic"
library(purrr)
library(here)
library(ggplot2)
library(dplyr)
library(scales)

set.seed(pi)

flip_cards <- readRDS(here("2026", "data", "flip-cards-function.RDS"))

psychic_scores <- seq(0, 1, 0.01)
success_scores <- map(psychic_scores, \(x) flip_cards(10, 4e6, x))


success_ratio <- function(x, threshold){
  sum(x >= threshold) / length(x)
}

scores <- list(
  anon_1 = 26/11 * 10,
  JenniK = 17.333,
  MagicIanthemagician = 190.666/52 * 10
)

score_prob <- map_dbl(success_scores, success_ratio, 82.3)
score_prob_anon_1 <- map_dbl(success_scores, success_ratio, scores$anon_1)
score_prob_JenniK <- map_dbl(success_scores, success_ratio, scores$JenniK)
score_prob_magician <- map_dbl(success_scores, success_ratio, scores$MagicIanthemagician)
score_prob_dan <- map_dbl(success_scores, success_ratio, 4.33)

len <- length(score_prob)

scores <- tibble(
  
  subject = 
    rep(
    c("Alex", "Anon", "JenniK", "MagicIanthemagician", "Dan Palmer"),
    each = len
  ),
  
  score_prob = c(score_prob, score_prob_anon_1, score_prob_JenniK, score_prob_magician, score_prob_dan)
)



results <-
  group_by(scores, subject) |> 
  mutate(psych_score = psychic_scores,
         norm_score_prob = (score_prob - min(score_prob)) / max(score_prob),
         norm_prob_diff = norm_score_prob - lag(norm_score_prob, default = 0)) 

results |> 
  filter(subject == "Alex") |> 
  
  ggplot(aes(x = psych_score, y = score_prob)) +
  
  geom_area(fill = "blue", colour = "black", alpha = 0.5) +
  
  labs(x = "Psychic score", y = "% beating 82.3", title = "Subject: Alex Dobson") +
  
  scale_y_continuous(labels = label_percent(1)) +
  
  theme_minimal()
```




```{r plotto}


ggplot(results) + 
  
  geom_area(aes(x = psych_score, y = norm_prob_diff, fill = subject),
            alpha = 0.5, position = "dodge", colour = "black") + 
  
  theme_minimal() +
  
  labs(
    x = "Psychic score",
    y = "Probability density"
  )

```

## The Puzzle {#sec-puzzle}

Not actually a puzzle at all. It's your turn to test yourself for psychic powers, using this [clever widget](https://church-army.shinyapps.io/psychic-tester/) I've built. Tune in next month to find out whether or not you are psychic!
